---
title: "bookdown: Writing Books with R Markdown"
output:
  html_document:
    self_contained: no
    mathjax: null
    theme: null
    highlight: null
    includes:
      in_header: in_header.html
      before_body: _navbar.html
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
sys.source('sitemap.R', globalenv())
```

<div class="container-fluid main-container">

<div class="tab-content">

<div id="home" class="tab-pane active">

```{r, echo=FALSE, message=FALSE}
book_panels()
```

</div>

<div id="get-started" class="tab-pane">

If this is the first time that you have ever heard of **bookdown**, you may get started with the minimal example [bookdown-demo](https://github.com/rstudio/bookdown-demo) on Github. Here are some simples steps to go through:

1. Fork or clone the repository <https://github.com/rstudio/bookdown-demo> if you are familiar with GIT and Github, or just download it as a Zip file then unzip it locally;
1. Install the [Preview version](https://www.rstudio.com/products/rstudio/download/preview/) of RStudio IDE (>= v0.99.1120);
1. Open the `bookdown-demo` repository you cloned or downloaded in RStudio by clicking `bookdown-demo.Rproj`;
1. Install the R package **bookdown**:

    ```{r eval=FALSE}
    if (!requireNamespace('devtools')) install.packages('devtools')
    devtools::install_github('rstudio/bookdown')
    ```

1. Open the R Markdown file `index.Rmd` and click the button `Knit` on the toolbar of RStudio, or click the `Build Book` button in the `Build` tab if you want to render the book to a specific format or all formats;

Now you should see the index page of this book demo in the RStudio Viewer (or the preview window). You may add or change the R Markdown files, come back to `index.Rmd`, and hit the `Knit` button again to preview the book. If you prefer not to use RStudio, you may also compile the book through command line. For the comprehensive documentation of **bookdown**, please see https://bookdown.org/yihui/bookdown.

To publish your books to bookdown.org, you need to first create an account by [signing in](/connect/) with your Google account, and then call the function `publish_book()` in R:

```{r eval=FALSE}
bookdown::publish_book(render = 'local')
```

If it is the first time you have tried to publish the book, you will be asked to authorize **bookdown** to publish to your bookdown.org account.

You are welcome to report any issues you encounter during writing your book (https://github.com/rstudio/bookdown/issues), and we hope you will enjoy **bookdown**.

</div> 

<div id="about" class="tab-pane">

The [**bookdown**](https://github.com/rstudio/bookdown) package is a free and open-source R package built on top of [R Markdown](http://rmarkdown.rstudio.com) to make it really easy to write books and long-form articles/reports. Markdown is a very simple language but made powerful thanks to [Pandoc](http://pandoc.org), and **bookdown** has added a few important missing features related to writing books, such as figure/table caption numbering and cross-references, and embedding [HTML widgets](https://htmlwidgets.org) or [Shiny apps](https://shiny.rstudio.com). We have tried hard to make everything work for all output formats (PDF, HTML, and EPUB, etc), so your readers can choose their favorite file format to read. Although the **bookdown** package was developed using R, it does not mean your book have to be related to R at all. You can certainly write poems or novels with **bookdown**!

The website bookdown.org is a service provided by [RStudio Inc.](https://www.rstudio.com) to host books. It is free for you to publish the static output files of your book, and you hold the full copyright of your own books.

</div> 

---

<footer class="footer text-center">
&copy; Copyright 2016 [RStudio Inc.](https://www.rstudio.com) | All Rights Reserved |
[<i class="fa fa-twitter"></i>](https://twitter.com/rstudio)
[<i class="fa fa-github"></i>](https://github.com/rstudio)
[<i class="fa fa-linkedin"></i>](https://www.linkedin.com/company/rstudio-inc)
[<i class="fa fa-facebook"></i>](https://www.facebook.com/pages/RStudio-Inc/267733656584415)
</footer>

</div>

</div>
