
Getting Started
=================================

## Install {.bigNumber .color1}

Install the [Preview version](https://www.rstudio.com/products/rstudio/download/preview/) of RStudio IDE (>= v0.99.1121);

Install the R package **bookdown**:

```{r eval=FALSE}
install.packages('devtools')
devtools::install_github('rstudio/bookdown')
```

## Edit {.bigNumber .color2}

Fork or clone the repository <https://github.com/rstudio/bookdown-demo> if you are familiar with GIT and Github, or just download it as a Zip file then unzip it locally;

Open the R Markdown file `index.Rmd` and click the button `Knit` on the toolbar of RStudio, or click the `Build Book` button in the `Build` tab if you want to render the book to a specific format or all formats;

Now you should see the index page of this book demo in the RStudio Viewer (or the preview window). You may add or change the R Markdown files, come back to `index.Rmd`, and hit the `Knit` button again to preview the book. If you prefer not to use RStudio, you may also compile the book through command line. For the comprehensive documentation of **bookdown**, please see https://bookdown.org/yihui/bookdown.

## Publish {.bigNumber .color2}

To publish your books to bookdown.org, you need to first create an account by [signing in](/connect/) with your Google account, and then call the function `publish_book()` in R:

```{r eval=FALSE}
bookdown::publish_book(render = 'local')
```

If it is the first time you have tried to publish the book, you will be asked to authorize **bookdown** to publish to your bookdown.org account.






