# Introduction {#intro}

All chapters start with a first-level heading followed by your chapter title, like the line above.

## Chapter sections {#a-section}

All chapter sections start with a second-level or higher heading followed by your section title, like the sections above and below.

## Heading labels

You can add a label to any heading using `{#nice-label}`, where `nice-label` is a placeholder. This chapter and the first section inside it are both labeled.

Heading labels let you reference a chapter or section from text elsewhere in your book with a clickable jump to that heading: Please review Chapter \@ref(intro). 

If you do not add manual labels (like in the current section), automatically valid labels are created based on the chapter name: See \@ref(heading-labels).

If you prefer text instead of numbered references, you can customize the link text: [any text you want can go here](#methods).

## Figures, tables, and captions

Like heading labels, figures and tables *with captions* can also be cross-referenced from elsewhere in your book using the `figure` and `table` environments, respectively.

```{r nice-fig, fig.cap='Here is a nice figure!', out.width='80%', fig.asp=.75, fig.align='center', fig.alt='Plot with connected points showing that vapor pressure of mercury increases exponentially as temperature increases.'}
par(mar = c(4, 4, .1, .1))
plot(pressure, type = 'b', pch = 19)
```

Reference a captioned figure using Figure \@ref(fig:nice-fig) in your text, where `fig:` is the required prefix and `nice-fig` is a placeholder for any code chunk label you want that produces a figure.

Similarly, you can reference tables generated from `knitr::kable()` using Table \@ref(tab:nice-tab), where `tab:` is the required prefix.

```{r nice-tab, tidy=FALSE}
knitr::kable(
  head(pressure, 10), caption = 'Here is a nice table!',
  booktabs = TRUE
)
```

## Citations

You can write citations, too. For example, we are using the **bookdown** package [@R-bookdown] (check out the last code chunk in index.Rmd to see how this citation key was added) in this sample book, which was built on top of R Markdown and **knitr** [@xie2015] (this citation was added in the external file book.bib).
