# Hello cross-references {#cross}

## Chapters and sub-chapters

Label a heading: `# Hello world {#nice-label}`

Reference a labeled heading in text: use `\@ref(nice-label)`; for example, please see Chapter \@ref(cross). 

Reference an unlabeled heading in text (`# Hello bookdown`): use `\@ref(hello-bookdown)`; for example, see Chapter \@ref(hello-bookdown). 

To un-number *and* label a heading: use `# Hello world {-#nice-label}`

Use text instead of a numbered reference: [any text you want can go here](#cross).

## Captioned figures and tables

Figures *with captions* can also be cross-referenced from elsewhere in your book using `\@ref(fig:chunk-label)` and `\@ref(tag:chunk-label)`, respectively.

```{r nice-fig, fig.cap='Here is a nice figure!', out.width='80%', fig.asp=.75, fig.align='center', fig.alt='Plot with connected points showing that vapor pressure of mercury increases exponentially as temperature increases.'}
par(mar = c(4, 4, .1, .1))
plot(pressure, type = 'b', pch = 19)
```

See Figure \@ref(fig:nice-fig).

```{r nice-tab, tidy=FALSE}
knitr::kable(
  head(pressure, 10), caption = 'Here is a nice table!',
  booktabs = TRUE
)
```

Don't miss Table \@ref(tab:nice-tab).
