--- 
yaml: goes here
---

# About

This is a _sample_ book written in **Markdown**. You can use anything that Pandoc's Markdown supports; for example, a math equation $a^2 + b^2 = c^2$.

## Installation

The **bookdown** package can be installed from CRAN or Github:

```{r eval=FALSE}
install.packages("bookdown")
# or the development version
# devtools::install_github("rstudio/bookdown")
```

## Usage 

Each chapter is an .Rmd file, and each .Rmd file can contain one (and only one) chapter. Each chapter *must* start with a first-level heading: `# A good chapter`.

You'll want to decide on at least one **bookdown** output format - your format(s) of choice go in the `_output.yml` file (you may have more than one). Here are the possible format options:
https://pkgs.rstudio.com/bookdown/reference/index.html#section-book-output-formats

You can build the HTML version of this example book without changing anything:

1. Find the **Build** pane in the RStudio IDE, and

1. Click on **Build Book**, then select your output format, or select "All formats" if you'd like to use multiple formats from the same book source files.

You can also build the book from the R console:

```{r, eval=FALSE}
bookdown::render_book()
```

You can also start a local server to live preview this HTML book. This preview will update as you edit the book when you save individual .Rmd files. You can start the server in a work session by using the RStudio add-in "Preview book", or from the R console:

```{r eval=FALSE}
bookdown::serve_book()
```

To compile this example to PDF as a `bookdown::pdf_book`, you'll need to install XeLaTeX. You are recommended to install TinyTeX (which includes XeLaTeX): <https://yihui.org/tinytex/>.

## Resources

The **bookdown** book: https://bookdown.org/yihui/bookdown/

The **bookdown** package reference site: https://pkgs.rstudio.com/bookdown


```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```
