# Components

## Markdown syntax

We give a very brief introduction to Pandoc's Markdown in this section. Readers who are familiar with Markdown can skip this section. The comphrehensive syntax of Pandoc's Markdown can be found on the Pandoc website <http://pandoc.org>.

You can make text _italic_ by surrounding it with underscores or asterisks, e.g., `_text_` or `*text*`. For **bold** text, use two underscores (`__text__`) or asterisks (`**text**`). To mark text as `inline code`, use a pair of backticks, e.g., `` `code` ``^[To include literal backticks, just use more backticks outside, e.g., you can use two backticks to preserve one backtick inside: ``` `` `code` `` ```.]. Links are created using `[text](link)`, e.g., `[RStudio](http://www.rstudio.com)`, and the syntax for images is similar: just add an exclamation mark, e.g., `![alt text or image title](path/to/image)`. Footnotes are put inside the square brackets after a caret `^[]`, e.g., `^[This is a footnote.]`. We will talk about citations in Section \@ref(citations). Section headers can be written after a number of pound signs, e.g.,

```
# First-level header

## Second-level header

### Third-level header
```

Unordered list items start with `*`, `-`, or `+`, and you can nest one list within another list by indenting the sub-list by four spaces, e.g.,

```
- one item
- one item
- one item
    - one item
    - one item
```

The output is:

- one item
- one item
- one item
    - one item
    - one item

Ordered list items start with numbers (the rule for nested lists is the same as above), e.g.,

```
1. the first item
2. the second item
3. the third item
```

The output does not look too much different with the Markdown source:

1. the first item
2. the second item
3. the third item

Blockquotes are written after `>`, e.g.,

```
> "I was always taught to respect my elders and
  I've now reached the age when I don't have anybody
  to respect", said George Burns.
```

The actual output:

> "I was always taught to respect my elders and
  I've now reached the age when I don't have anybody
  to respect", said George Burns.

Plain code blocks can be written after three or more backticks, and you can also indent the blocks by four spaces, e.g.,

````
```
This text is displayed verbatim / preformatted
```

Or indent by four spaces:

    This text is displayed verbatim / preformatted
````

Inline LaTeX equations can be written in a pair of dollar signs using the LaTeX syntax, e.g., `$f(k) = {n \choose k} p^{k} (1-p)^{n-k}$` (actual output: $f(k)={n \choose k}p^{k}(1-p)^{n-k}$); math expressions of the display style can be written in a pair of double dollar signs, e.g., `$$f(k) = {n \choose k} p^{k} (1-p)^{n-k}$$`, and the output looks like this:

$$f(k)={n \choose k}p^{k}(1-p)^{n-k}$$

## R code

There are two types of R code in R Markdown/**knitr** documents: R code chunks, and inline R code. The syntax for the latter is `` ``r ''`r R_CODE` ``, and it can be embedded inline with other document elements. R code chunks look like plain code blocks, but has `{r}` after the three backticks and (optionally) chunk options inside `{}`, e.g.,

    `r ''````{r chunk-label, echo = FALSE, fig.cap = 'A figure caption.'}
    1 + 1
    rnorm(10)  # 10 random numbers
    plot(dist ~ speed, cars)  # a scatterplot
    ```

To learn more about **knitr** chunk options, see @xie2015 or <http://yihui.name/knitr/options>.

## Figures {#figures}

By default, figures have no captions in the output generated by **knitr**, which means they will just be placed wherever they were generated in the R code. Below is such an example.

```{r no-caption, fig.width=6, fig.asp=.7, out.width='70%', ref.label='pressure-plot'}
```

The disadvantage of typesetting figures in this way is that when there is not enough space on the current page to place a figure, it may either reach the bottom of the page (hence exceeds the page margin), or be pushed to the next page, leaving a large white margin at the bottom of the current page. That is basically why there are "floating environments" in LaTeX: elements that cannot be split over multiple pages (like figures) are put in floating environments, so they can float to a page that has enough space to hold them. There is also a disadvantage of floating things forward or backward, though. That is, readers may have to jump to a different page to find the figure mentioned on the current page. This is simply a natural consequence of having to typeset things on multiple pages of fixed sizes. This issue does not exist in HTML, however, since everything can be placed continuously on one single page (presumably with infinite height), and there is no need to split anything across multiple pages of the same page size.

If we assign a figure caption to a code chunk via the chunk option `fig.cap`, R plots will be put into figure environments, which will be automatically labeled and numbered, and can also be cross-referenced. The label of a figure environment is generated from the label of the code chunk, e.g., if the chunk label is `foo`, the figure label will be `fig:foo` (the prefix `fig:` is added before `foo`). To reference a figure, use the syntax `\@ref(label)`^[Do not forget the leading backslash!], where `label` is the figure label, e.g., `fig:foo`.

```{block, type='rmdimportant'}
If you want to cross-reference figures or tables generated from a code chunk, please make sure the chunk label only contains alphanumeric characters (a-z, A-Z, 0-9) and dashes (-). Other characters do not qualify.
```


The chunk option `fig.asp` can be used to set the aspect ratio of plots, i.e., the ratio of figure height/width. If the figure width is 6 inches (`fig.width = 6`) and `fig.asp = 0.7`, the figure height will be automatically calculated from `fig.width * fig.asp = 6 * 0.7 = 4.2`. Figure \@ref(fig:pressure-plot) is an example using the chunk options `fig.asp = 0.7`, `fig.width = 6`, and `fig.align = 'center'`, generated from the code below:

```{r pressure-plot, fig.asp=.7, fig.width=6, fig.cap='A figure example with the specified aspect ratio, width, and alignment.', fig.align='center', out.width='80%'}
par(mar = c(4, 4, .1, .1))
plot(pressure, pch = 19, type = 'b')
```

The actual size of a plot is determined by the chunk options `fig.width` and `fig.height` (the size of the plot generated from a graphical device), and we can specify the output size of plots via the chunk options `out.width` and `out.height`. The possible value of these two options depends on the output format of the document. For example, `out.width = '30%'` is a valid value for HTML output, but not for LaTeX/PDF output. However, **knitr** will automatically convert a percentage value for `out.width` of the form `x%` to `(x / 100) \linewidth`, e.g., `out.width = '70%'` will be treated as `.7\linewidth` when the output format is LaTeX. This makes it possible to specify a relative width of a plot in a consistent manner. Figure \@ref(fig:cars-plot) is an example of `out.width = 70%`.

```{r cars-plot, out.width='70%', fig.cap='A figure example with a relative width 70\\%.'}
par(mar = c(4, 4, .1, .1))
plot(cars, pch = 19)
```

Sometimes you may want to put multiple plots in one figure environment. In this case, you must use the chunk option `fig.show = 'hold'` to hold multiple plots from a code chunk and include them in one environment. You can also place plots side by side if the sum of the width of all plots is smaller than or equal to the current line width. For example, if two plots have the same width `50%`, they will be placed side by side. Similarly, you can specify `out.width = '33%'` to arrange three plots on one line. Figure \@ref(fig:multi-plots) is an example of two plots, each with a width `50%`.

```{r multi-plots, out.width='50%', fig.show='hold', fig.cap='Two plots placed side by side.'}
par(mar = c(4, 4, .1, .1))
plot(pressure, pch = 19, type = 'b')
plot(cars, pch = 19)
```



## Tables {#tables}

For now, the most convenient way to generate a table is the function `knitr::kable()`, because there are some internal tricks in **knitr** to make it work with **bookdown** and users do not have to know anything about these implementation details. We will explain how to use other packages and functions later in this section.

Like figures, tables with captions will also be numbered and can be referenced. The `kable()` function will automatically generate a label for a table environment, which is the prefix `tab:` plus the chunk label. For example, the table label for a code chunk with the label `foo` will be `tab:foo`, and we can still use the syntax `\@ref(label)` to reference the table. Table \@ref(tab:table-single) is a simple example.

```{r table-single, tidy=FALSE}
knitr::kable(
  head(mtcars, 10), booktabs = TRUE,
  caption = 'A table of the first 10 rows of the mtcars data.'
)
```

If you want to put multiple tables in a single table environment, just wrap the data objects (usually data frames in R) into a list. See Table \@ref(tab:table-multi) for an example.

```{r table-multi, tidy=FALSE}
knitr::kable(
  list(
    head(iris[,1:2],3),
    head(mtcars[,1:3],5)
  ),
  caption = 'A Tale of Two Tables.', booktabs = TRUE
)
```

If you decide to use other packages to generate tables, you have to make sure the label for the table environment appears in the beginning of the table caption in the form `(\#label)`, where `label` must have the prefix `tab:`. You have to be very careful about the _portability_ of the table generating function: it should work for both HTML and LaTeX output automatically, so it must consider the output format internally (check `knitr::opts_knit$get('pandoc.to')`). When writing out an HTML table, the caption must be written in the `<caption></caption>` tag. For simple tables, `kable()` should suffice. If you have to create complicated tables (e.g., with certain cells spanning across multiple columns/rows), you will have to take the aforementioned issues into consideration.

## Cross-references

We have explained how cross-references work for figures (Section \@ref(figures)) and tables (Section \@ref(tables)). In fact, you can also reference sections using the same syntax `\@ref(label)`, where `label` is the section ID. By default, Pandoc will generate an ID for all section headers, e.g., a section `# Hello World` will have an ID `hello-world`. We recommend you to manually assign an ID to a section header to make sure you do not forget to update the reference label after you change the section header. To assign an ID to a section header, simply add `{#id}` to the end of the section header.

When a referenced label cannot be found, you will see two question marks like \@ref(fig:does-not-exist), as well as a warning message in the R console when rendering the book.

## Custom blocks

You can generate custom blocks using the `block` engine in **knitr**, i.e., the chunk option `engine = 'block'`, or the more compact syntax ```` ```{block} ````. This engine should be used in conjunction with the chunk option `type`, which takes a character string. When the `block` engine is used, it generates a `<div>` to wrap the chunk content if the output format is HTML, and a LaTeX environment if the output is LaTeX. The `type` option specifies the class of the `<div>` and the name of the LaTeX environment. For example, the HTML output of this chunk

    `r ''````{block, type='FOO'}
    Some text for this block.
    ```

will be this:

    <div class="FOO">
    Some text for this block.
    </div>

and the LaTeX output will be this:

    \begin{FOO}
    Some text for this block.
    \end{FOO}

It is up to the book author how to define the style of the block. You can define the style of the `<div>` in CSS and include it in the output via the `includes` option in the YAML metadata. Similarly, you may define the LaTeX environment via `\newenvironment` and include the definition in the LaTeX output via the `includes` option. For example, we may save the following style in a CSS file, say, `style.css`:

```css
div.FOO {
  font-weight: bold;
  color: red;
}
```

And the YAML metadata of the R Markdown document can be:

```yaml
---
output:
  bookdown::html_chapters:
    includes:
      in_header: style.css
---
```

We have defined a few types of blocks for this book to show notes, tips, and warnings, etc. Below are some examples:

```{block, note-text, type='rmdnote'}
R is free software and comes with ABSOLUTELY NO WARRANTY.
You are welcome to redistribute it under the terms of the
GNU General Public License versions 2 or 3.
For more information about these matters see
http://www.gnu.org/licenses/.
```

```{block, note-text, type='rmdcaution'}
```


```{block, note-text, type='rmdimportant'}
```

```{block, note-text, type='rmdtip'}
```

```{block, note-text, type='rmdwarning'}
```

## Citations {#citations}

## Interactive Shiny apps

