# Publishing

As you develop the book, you may put the draft book in the public to get early feedback from readers, e.g., publish it to a website. After you finish writing the book, you need to think about options to formally publish it as either printed copies or e-books.

## RStudio Connect

In theory, you can render the book by yourself and publish the output anywhere you want. For example, you can host the HTML files on your own web server. We have provided a function `publish_book()` in **bookdown** to make it very simple to upload your book to https://bookdown.org, which is a website provided by RStudio to host your books for free. This website is built on top of "RStudio Connect", an RStudio product that allows you to deploy a variety of R-related applications to a server, including R Markdown documents, Shiny applications, R plots, and so on.

You do not have to know much about RStudio Connect to publish your book to bookdown.org. Basically you just sign up at https://bookdown.org/connect/, and the first time you try to run `bookdown::publish_book()`, you will be asked to authorize **bookdown** to publish to your bookdown.org account. In the future, you simply call `publish_book()` again and **bookdown** will no longer ask for anything.

```{r publish-book-usage, eval=FALSE, code=formatR::usage(bookdown::publish_book, output=FALSE)}
```

The only argument of `publish_book()` that you may want to touch is `render`. It determines whether you want to render the book before publishing. If you have run `render_book()` before, you do not need to change this argument, otherwise you may set it to `'local'`:

```{r eval=FALSE}
bookdown::publish_book(render = 'local')
````

If you have set up your own RStudio Connect server, you can certainly publish the book to that server instead of bookdown.org.

## GitHub

You can host your book on GitHub for free via GitHub Pages (https://pages.github.com). GitHub supports Jekyll (http://jekyllrb.com), a static website builder, to build a website from Markdown files. That may be the more common use case of GitHub Pages, but GitHub also supports arbitrary static HTML files, so you can just host the HTML output files of your book on GitHub. To publish to your book to GitHub Pages, you need to create a `gh-pages` branch in your repository, build the book, put the HTML output (including all external resources like images, CSS, and JavaScript files) in this branch, and push the branch to the remote repository.

If your book repository does not have the `gh-pages` branch, you may use the following commands to create one:

```bash
# assume you have initialized the git repository,
# and are under the directory of the book repository now

# create a branch named gh-pages and clean up everything
git checkout --orphan gh-pages
git rm -rf .

# create a hidden file .nojekyll
touch .nojekyll
git add .nojekyll

git commit -m"Initial commit"
git push origin gh-pages
```

The hidden file `.nojekyll` tells GitHub that your website is not to be built via Jekyll, since the **bookdown** HTML output is already a standalone website. If you are on Windows, you may not have the `touch` command, and you can just create the file in R using `file.create('.nojekyll')`.

After you have set up GIT, the rest of work can be automated via a script (Shell, R, or Makefile, depending on your preference). Basically, you compile the book to HTML, then run git commands to push the files to GitHub, but you probably do not want to do this over and over again manually and locally. It can be very handy to automate the publishing process completely on the cloud, so once it is set up correctly, all you have to do next is write the book and push the Rmd source files to GitHub, and your book will always be automatically built and published from the server side.

One service that you can utilize is Travis CI (https://travis-ci.org). It is free for public repositories on GitHub, and was designed for continuous integration (CI) of software packages. Travis CI can be connected to GitHub in the sense that whenever you push to GitHub, Travis can be triggered to run certain commands/scripts on the latest version of your repository^[You need to authorize the Travis CI service for your repository on GitHub first. See https://docs.travis-ci.com/user/getting-started/ for how to get started with Travis CI.]. These commands are specified in a YAML file named `.travis.yml` in the root directory of your repository, and they are usually for the purpose of testing software, but in fact they are quite open-ended, meaning that you can run arbitrary commands on a Travis (virtual) machine. That means you can certainly run your own scripts to build your book on Travis. Note Travis only supports Ubuntu and Mac OS X at the moment, so you should have some basic knowledge about Linux/Unix commands.

The next question is, how to publish the book built on Travis to GitHub? Basically you have to grant Travis write access to your GitHub repository. This authorization can be done via several ways, and the easiest one to beginners may be a personal access token. Below are a few steps you may follow:

1. Create a personal access token for your account on GitHub: https://help.github.com/articles/creating-an-access-token-for-command-line-use/ (make sure to enable the "repo" scope so that using this token will enable writing to your GitHub repos)
1. Encrypt it in the environment variable `GH_TOKEN` via command line `travis encrypt` and store it in `.travis.yml`. If you do not know how to install or use the Travis command-line tool, simply save this environment variable via `https://travis-ci.org/user/repo/settings` where `user` is your GitHub ID, and `repo` is the name of the repository;
1. You can clone this `gh-pages` branch on Travis using your GitHub token, add the HTML output files from R Markdown (do not forget to add figures and CSS style files as well), and push to the remote repository.

Assume you are in the `master` branch right now (where you put the Rmd source files), and have compiled the book to the `_book` directory. What you can do next on Travis is:

```bash
# configure your name and email if you have not done so
git config --global user.email "you@example.com"
git config --global user.name "Your Name"

# clone the repository to the book-output directory
git clone -b gh-pages \
  https://${GH_TOKEN}@github.com/${TRAVIS_REPO_SLUG}.git \
  book-output
cd book-output
cp -r ../_book/* ./
git add *
git commit -m"Update the book"
git push origin gh-pages
```

The variable name `GH_TOKEN` and the directory name `book-output` are arbitrary, and you can use any names you prefer, as long as the names do not conflict with existing environment variable names or directory names. This script, together with the build script we mentioned in Section \@ref(build-the-book), can be put in the `master` branch as Shell scripts, e.g., you can name them as `_build.sh` and `_deploy.sh`. Then your `.travis.yml` may look like this:

```yaml
language: r

env:
  global:
    - secure: A_LONG_ENCRYPTED_STRING

r_github_packages:
  - rstudio/bookdown

install: true

before_script:
  - chmod +x ./_build.sh
  - chmod +x ./_deploy.sh

script:
  - ./_build.sh
  - ./_deploy.sh
```

The `language` key tells Travis to use a virtual machine that has R installed. The `secure` key is your encrypted personal access token. If you have already saved the `GH_TOKEN` variable using the web interface on Travis instead of the command-line tool `travis encrypt`, you can leave out this key.

In the `install` step, the above example did not do anything, but you may want to install some dependencies for your book there, e.g.,

```yaml
install:
  - Rscript -e "if (!require('ggplot2')) install.packages('ggplot2')"
```

If you use the [container-based infrastructure](https://docs.travis-ci.com/user/workers/container-based-infrastructure/) on Travis, you can enable caching by using `sudo: false` in `.travis.yml`. Normally you should cache at least two types of directories: the figure directory (e.g., `_main_files`) and the cache directory (e.g., `_main_cache`). These directory names may also be different if you have specified the **knitr** chunk options `fig.path` and `cache.path`, but I'd strongly recommend you not to change these options. The figure and cache directories are stored under the `_bookdown_files` directory of the book root directory. A `.travis.yml` file that has enabled caching of **knitr** figure and cache directories may have additional configurations `sudo` and `cache` like this:

```yaml
sudo: false

cache:
  packages: yes
  directories:
    - $TRAVIS_BUILD_DIR/_bookdown_files
```

If your book is very time-consuming to build, you may use the above configurations on Travis to save time. Note `packages: yes` means the R packages installed on Travis are also cached.

GitHub and Travis CI are certainly not the only choices to build and publish your book. You are free to store and publish the book on your own server.

## Publishers

## Licensing

## Self-publishing
