--- 
title: "Authoring A Book with R Markdown"
author: "Yihui Xie"
date: "`r Sys.Date()`"
knit: "bookdown::render_book"
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
lapply(c('DT'), function(pkg) {
  if (!requireNamespace(pkg)) install.packages(pkg)
})
if (packageVersion('DT') < '0.1.45') devtools::install_github('rstudio/DT')
```

# Introduction

```{block, type='rmdwarning'}
This book and the package **bookdown** are still under active development, and should not be considered stable at the moment. You are welcome to experiment with this package, and feedback may be sent to https://github.com/rstudio/bookdown/issues or yihui@rstudio.com.
```

This book is a guide to authoring books with R Markdown [@R-rmarkdown] and the R package **bookdown** [@R-bookdown]. It focuses on the features specific to writing books, long-form articles, or reports, such as

- How to typeset figures and tables, and cross-reference them;
- How to generate multiple output formats such as HTML, PDF, and e-books for a single book;
- How to customize the book templates and style different elements in a book;
- The editor support (in particular, the RStudio IDE);
- How to publish a book;

It is not a comprehensive introduction to R Markdown or the **knitr** package [@R-knitr], on top of which **bookdown** was built. To learn more about R Markdown, please check out the online documentation <http://rmarkdown.rstudio.com>. For **knitr**, please see @xie2015. You do not have to be an expert of the R language [@R-base] to read this book, but you are expected to have some basic knowledge about R Markdown and **knitr**. For beginners, you may get started with the cheatsheets at <https://www.rstudio.com/resources/cheatsheets/>. To be able to customize the book templates and themes, you should be familiar with LaTeX, HTML and CSS.

## Motivation

Markdown is a wonderful language to write relatively simple documents that contain elements like sections, paragraphs, lists, links, and images, etc. Pandoc (<http://pandoc.org>) has greatly extended the original Markdown syntax, and added quite a few useful new features, such as footnotes, citations, and tables. More importantly, Pandoc makes it possible to generate output documents of a large variety of formats from Markdown, including HTML, LaTeX/PDF, Word, and slides.

To write a relatively complicated document like a book, there are still a few useful features missing in Pandoc's Markdown at the moment, such as automatic numbering of figures and tables in the HTML output, cross-references of figures and tables, and fine control of the appearance of figures (e.g., currently it is impossible to specify the alignment of images using the Markdown syntax). These are some of the problems that we have addressed in the **bookdown** package.

Under the constraint that we want to produce the book in multiple output formats, it is nearly impossible to cover all possible features specific to these output formats. For example, it may be difficult to reinvent a certain complicated LaTeX environment in the HTML output using the (R) Markdown syntax. Our main goal is not to replace _everything_ with Markdown, but to cover _most_ common functionalities required to write a relatively complicated document, and make the syntax of such functionalities consistent across all output formats, so that you only need to learn one thing and it works for all output formats.

Another goal of this project is to make it easy to produce books that look visually pleasant. Some nice existing examples include Gitbook (<https://www.gitbook.com>), Tufte CSS (<http://edwardtufte.github.io/tufte-css/>), and Tufte-LaTeX (<https://tufte-latex.github.io/tufte-latex/>). We hope to integrate these themes and styles into **bookdown**, so authors do not have to dive into the details of how to use a certain LaTeX class or how to configure CSS for HTML output.

## Get started

The easiest way for beginners to get started with writing a book with R Markdown and **bookdown** is through the demo `bookdown-demo` on GitHub:

1. Fork or clone the GitHub repository <https://github.com/rstudio/bookdown-demo> if you are familiar with GIT and GitHub, or just download it as a Zip file then unzip it locally;
1. Install the RStudio IDE (<http://www.rstudio.com>) if you have not done so;
1. Open the `bookdown-demo` repository you cloned or downloaded in RStudio by clicking `bookdown-demo.Rproj`;
1. Install the R package **bookdown**:

    ```{r eval=FALSE}
    devtools::install_github('rstudio/bookdown')
    ```

1. Open the R Markdown file `index.Rmd` and click the button `Knit` on the toolbar of RStudio;

Now you should see the index page of this book demo in the RStudio Viewer. You may add or change the R Markdown files, come back to `index.Rmd`, and hit the `Knit` button again to preview the book. If you prefer not to use RStudio, you may also compile the book through command line. See the next section for details.

## Usage

Normally, a book contains multiple chapters, and one chapter lives in one R Markdown file, with the filename extension `.Rmd`. Each R Markdown file must start immediately with the chapter title. All R Markdown files must be encoded in UTF-8. Here is an example (the bullets are the filenames, followed by the file content):

- 01-intro.Rmd

    ```markdown
    # Introduction
    
    This chapter is an overview of the methods that
    we propose to solve an **important problem**.
    ```

- 02-literature.Rmd

    ```markdown
    # Literature
    
    Here is a review of existing methods.
    ```

- 03-method.Rmd

    ```markdown
    # Methods
    
    We describe our methods in this chapter.
    ```

- 04-application.Rmd

    ```markdown
    # Applications
    
    Some _significant_ applications are demonstrated
    in this chapter.
    
    ## Example one
    
    ## Example two
    ```

- 05-summary.Rmd

    ```markdown
    # Final Words
    
    We have finished a nice book.
    ```

By default, **bookdown** merges all Rmd files by the order of filenames, e.g., `01-intro.Rmd` will appear before `02-literature.Rmd`. Filenames that start with an underscore `_` are skipped. If there exists an Rmd file named `index.Rmd`, it will always be treated as the first file when merging all Rmd files. The reason for this special treatment is that the HTML file `index.html` to be generated from `index.Rmd` is usually the default index file when you view a website, e.g., you are actually browsing http://yihui.name/index.html when you open http://yihui.name/.

You can override the above behavior by including a configuration file named `_bookdown.yml` in the book directory. It is a YAML file (https://en.wikipedia.org/wiki/YAML), and R Markdown users should be familiar with this format since it is also used to write the metadata in the beginning of R Markdown documents. You can use a field named `rmd_files` to define your own list and order of Rmd files for the book. For example,

```yaml
rmd_files: ["index.Rmd", "abstract.Rmd", "intro.Rmd"]
```

In this case, **bookdown** will just use whatever you defined in this YAML field without any special treatments of `index.Rmd` or underscores. If you want both HTML and LaTeX/PDF output from the book, and use different Rmd files for HTML and LaTeX output, you may specify these files for the two output formats separately, e.g.,

```yaml
rmd_files:
  html: ["index.Rmd", "abstract.Rmd", "intro.Rmd"]
  latex: ["abstract.Rmd", "intro.Rmd"]
```

```{block, type='rmdimportant'}
Because **knitr** does not allow duplicate chunk labels in a source document, you need to make sure there are no duplicate labels in your book chapters, otherwise **knitr** will signal an error when knitting the merged Rmd file. 
```

Although we have been talking about R Markdown files, the chapter files do not actually have to be R Markdown. They can be plain Markdown files (`.md`), and do not have to contain R code chunks at all. You can certainly use **bookdown** to compose novels or poems!

At the moment, there are three output formats that you may use: `bookdown::pdf_book`, `bookdown::gitbook`, and `bookdown::html_chapters`. There is a `bookdown::render_book()` function similar to `rmarkdown::render()`, but it was designed to render _multiple_ Rmd documents into a book using the output format functions. You may either call this function from command line, or use it in the RStudio IDE. Here are some command line examples:

```{r eval=FALSE}
bookdown::render_book('foo.Rmd', 'bookdown::gitbook')
bookdown::render_book('foo.Rmd', 'bookdown::pdf_book')
bookdown::render_book('foo.Rmd', bookdown::gitbook(lib_dir = 'book_assets'))
bookdown::render_book('foo.Rmd', bookdown::pdf_book(keep_tex = TRUE))
```

To use `render_book` and the output format functions in the RStudio IDE, you can define a YAML field named `knit` that takes the value `bookdown::render_book`, and the output format functions can be used in the `output` field, e.g.,

```yaml
---
knit: "bookdown::render_book"
output:
  bookdown::gitbook:
    lib_dir: "book_assets"
  bookdown::pdf_book:
    keep_tex: yes
---
```

Then you can click the `Knit` button in RStudio to compile the Rmd files into a book.

## Configuration

We have mentioned `rmd_files`, and there are more things you can configure for a book in `_bookdown.yml`:

- `book_filename`: the filename of the main Rmd file, i.e., the Rmd file that is merged from all chapters; by default, it is named `_main.Rmd`.
- `chapter_name`: (for HTML output only) either a character string to be prepended to the chapter number in the chapter title (e.g., `'Chapter '`), or an R function that takes the chapter number as the input and returns a string as the new chapter number (e.g., `!expr function(i) paste('Chapter', i)`).
- `before_chapter_script`: one or multiple R scripts to be executed before each chapter, e.g., you may want to clear the workspace before compiling each chapter, in which case you can use `rm(list = ls(all = TRUE))` in the R script.
- `after_chapter_script`: similar to `before_chapter_script`, and the R script is executed after each chapter.
- `edit`: a link that collaborators can click to edit the Rmd source document of the current page; this was designed primarily for Github repositories, since it is easy to edit arbitrary plain-text files on Github even in other people's repositories (if you do not have write access to the repository, Github will automatically fork it and let you submit a pull request after you finish editing the file). This link is should have `%s` in it, which will be substituted by the actual Rmd filename for each page.
    - Optionally, you can have a `text` field under the `edit` field to specify the text to which the link is attached.
- `output_dir`: the output directory of the book; this setting is read and used by `render_book()`.
- `clean`: a vector of files and directories to be cleaned by the `clean_book()` function.

Here is a sample `_bookdown.yml`:

```yaml
book_filename: "my-book.Rmd"
chapter_name: "CHAPTER "
before_chapter_script: ["script1.R", "script2.R"]
after_chapter_script: "script3.R"
edit:
  link: https://github.com/rstudio/bookdown/edit/master/inst/examples/%s
  text: "Edit"
output_dir: "book-output"
clean: ["my-book.bbl", "R-packages.bib"]
```

Besides the configurations in `_bookdown.yml`, you can also specify some Pandoc-related configurations in the YAML metadata of the _first_ Rmd file of the book, such as the title, author, and date of the book, etc. For example:

```yaml
--- 
title: "Authoring A Book with R Markdown"
author: "Yihui Xie"
date: "`r Sys.Date()`"
knit: "bookdown::render_book"
output:
  bookdown::gitbook: default
documentclass: book
bibliography: ["book.bib", "packages.bib"]
biblio-style: apalike
link-citations: yes
---
```

## Two rendering approaches {#new-session}

Merging all chapters into one Rmd file and knitting it is one way to render the book in **bookdown**. There is actually another way: you may knit each chapter in a _separate_ R session, and **bookdown** will merge the Markdown output of all chapters to render the book. We call these two approaches "Merge and Knit" (MK) and "Knit and Merge" (KM), respectively. The differences between them may seem subtle, but can be fairly important depending on your use cases.

- The most significant difference is that MK runs _all_ code chunks in all chapters in the same R session, whereas KM uses separate R sessions for individual chapters. For MK, the state of the R session from previous chapters is carried over to later chapters (e.g., objects created in previous chapters are available to later chapters, unless you deliberately deleted them); for KM, all chapters are isolated from each other^[Of course, no one can stop you from writing out some files in one chapter, and reading them in another chapter.]. If you want each chapter to compile from a clean state, use the KM approach. It can be very tricky and difficult to restore a running R session to a completely clean state if you use the MK approach. For example, even you detach/unload packages loaded in a previous chapter, R will not clean up the S3 methods registered by these packages.
- One advantage of KM is that Rmd files that have not been updated since the last time the book was rendered will not be recompiled by default, unless you force all chapters to be recompiled via `render_book(force_knit = TRUE)`. This may save some time, but the speedup may not be very significant, since the major time is normally consumed by running code chunks. If time-consumed chunks are cached, the compilation time for MK and KM may be about the same.
- The KM approach will generate more files under the directory of Rmd files: each Rmd file will generate a Markdown output file (`.md`), and possibly a figure directory and a cache directory (e.g. `01-intro_files/` and `01-intro_cache/`). The MK approach only renders one Rmd file, so it only has one set of output files.
- For KM, whenever you change the output format (e.g., from HTML to PDF), you must recompile all chapters (`render_book(force_knit = TRUE)`), because the Markdown output files for one format may not work for another format. There is no such issue with the MK approach.

The default approach in **bookdown** is MK. To switch to KM, you either use the argument `new_session = TRUE` when calling `render_book()`, or set `new_session: yes` in the configuration file `_bookdown.yml`.

You can still configure `book_filename` in `_bookdown.yml` for the KM approach, but it should be a Markdown filename, e.g., `_main.md`, although the filename extension does not really matter, and you can even leave out the extension, e.g., just set `book_filename: _main`. All other configurations work for both MK and KM.

## Some tips

Typesetting under the paging constraint (e.g., for LaTeX/PDF output) can be an extremely tedious and time-consuming job. I'd recommend you not to look at your PDF output frequently, since most of the time you are very unlikely to be satisfied: text may overflow into the page margin, figures may float too far away, and so on. Do not try to make things look right _immediately_, because you may be disappointed over and over again as you keep on revising the book, and things may be messed up again even if you only did some minor changes (see <http://bit.ly/tbrLtx> for a nice illustration).

If you want to preview the book, preview the HTML output. Work on the PDF book after you have finished the content of the book, and are very sure no major revisions will be required.

If certain code chunks in your R Markdown docuemnts are time-consuming to run, you may cache them by adding the chunk option `cache = TRUE` in the chunk header, and you are recommended to label such code chunks as well, e.g.,

    `r ''````{r important-computing, cache=TRUE}

We will talk about how to quickly preview books as you keep on editing it in Chapter \@ref(editors). In short, you can use the `preview_chapter()` function to render a single chapter instead of the whole book. The function `serve_book()` makes it easy to live-preview HTML book pages: whenever you modify an Rmd file, the book can be recompiled and the browser can be automatically refreshed accordingly.
